Inference methods comparison for the simplified Naomi model

### 2023-02-13 src/naomi-simple_compare/comparison.Rmd
Correct error with starts_with option, add green highlighting for lowest KS value in table, alter the summary plots to be more appropriate

# Histograms and empirical cumulative distribution plots {.tabset .tabset-fade}

We create histograms and empirical cumulative distribution function (ECDF) plots of the samples from each method.
All of the possible parameter names are as follows:

```{r}
pars <- unique(names(tmb$fit$obj$env$par))
```

@@

```{r}
ks_helper <- function(par, ...) {
  to_ks_df(par) %>% ks_plot(par, ...)
}
```

@@

```{r}
ks_helper("beta")
ks_helper("beta", method1 = "TMB", method2 = "adam")
```

@@

```{r}
ks_helper("logit")
ks_helper("logit", method1 = "TMB", method2 = "adam")
```

@@

```{r}
ks_helper("log_sigma")
ks_helper("log_sigma", method1 = "TMB", method2 = "adam")
```

@@

```{r}
ks_summary <- lapply(unique(names(tmb$fit$obj$env$par)), function(x) {
  to_ks_df(x) %>%
    group_by(method) %>%
    summarise(ks = mean(ks), par = x)
}) %>%
  bind_rows() %>%
  pivot_wider(names_from = "method", values_from = "ks") %>%
  rename(
    "Parameter" = "par",
    "KS(adam, tmbstan)" = "adam",
    "KS(aghq, tmbstan)" = "aghq",
    "KS(TMB, tmbstan)" = "TMB",
  )

ks_summary %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = starts_with("KS"),
    decimals = 3
  )

@@

dict <- data.frame(
  Parameter = c(x_names, theta_names),
  Type = c(rep("Latent field", length(x_names)), rep("Hyperparameters", length(theta_names)))
)

ks_summary <- ks_summary %>%
  left_join(dict, by = "Parameter")

summary_ks_plot <- function(ks_summary, method1, method2) {
  ks_method1 <- paste0("KS(", method1, ", tmbstan)")
  ks_method2 <- paste0("KS(", method2, ", tmbstan)")

  scatterplot <- ggplot(ks_summary, aes(x = .data[[ks_method1]], y = .data[[ks_method2]])) +
    geom_point(alpha = 0.2) +
    annotate(geom = "polygon", x = c(-Inf, Inf, Inf), y = c(-Inf, Inf, -Inf), fill = multi.utils::cbpalette()[3], alpha = 0.1) +
    annotate(geom = "polygon", x = c(-Inf, Inf, -Inf), y = c(-Inf, Inf, Inf), fill = multi.utils::cbpalette()[5], alpha = 0.1) +
    xlim(0, 1) +
    ylim(0, 1) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
    labs(x = ks_method1, y = ks_method2) +
    theme_minimal()

@@

  jitterplot <- ggplot(ks_summary, aes(x = Type, y = `KS difference`)) +
    geom_jitter(width = 0.1, alpha = 0.2) +
    labs(x = "", y = paste0(ks_method1, " - ", ks_method2)) +
    theme_minimal()

  scatterplot + jitterplot
}

>>>>

# Histograms and ECDF difference plots {.tabset .tabset-fade}

We create histograms and empirical cumulative distribution function (ECDF) difference plots of the samples from each method.
All of the possible parameter names are as follows:

```{r}
pars <- unique(names(tmb$fit$obj$env$par))
pars
```

@@

```{r}
ks_helper <- function(par, sw = FALSE, ...) {
  to_ks_df(par, starts_with = sw) %>% ks_plot(par, ...)
}
```

@@

```{r}
ks_helper("beta", sw = TRUE)
ks_helper("beta", sw = TRUE, method1 = "TMB", method2 = "TMB")
```

@@

```{r}
ks_helper("logit", sw = TRUE)
ks_helper("logit", sw = TRUE, method1 = "TMB", method2 = "adam")
```

@@

```{r}
ks_helper("log_sigma", sw = TRUE)
ks_helper("log_sigma", sw = TRUE , method1 = "TMB", method2 = "adam")
```

@@

```{r}
options(dplyr.summarise.inform = FALSE)

ks_summary <- lapply(unique(names(tmb$fit$obj$env$par)), function(x) {
  to_ks_df(x) %>%
    group_by(method, index) %>%
    summarise(ks = mean(ks), par = x) %>%
    ungroup()
}) %>%
  bind_rows() %>%
  pivot_wider(names_from = "method", values_from = "ks") %>%
  rename(
    "Parameter" = "par",
    "KS(adam, tmbstan)" = "adam",
    "KS(aghq, tmbstan)" = "aghq",
    "KS(TMB, tmbstan)" = "TMB",
  )

@@

dict <- data.frame(
  Parameter = c(unique(x_names), unique(theta_names)),
  Type = c(rep("Latent field", length(unique(x_names))), rep("Hyper", length(unique(theta_names))))
)

ks_summary <- ks_summary %>%
  left_join(dict, by = "Parameter")
```

Values in green are the lowest KS difference for that particular parameter:

```{r}
ks_summary %>%
  gt::gt() %>%
  gt::fmt_number(
    columns = starts_with("KS"),
    decimals = 3
  ) %>%
  gt::tab_style(
    style = cell_fill(color = "#009E73"),
    locations = cells_body(
      columns = `KS(adam, tmbstan)`,
      rows = `KS(adam, tmbstan)` < `KS(aghq, tmbstan)` & `KS(adam, tmbstan)` < `KS(TMB, tmbstan)`
    )
  ) %>%
    gt::tab_style(
    style = cell_fill(color = "#009E73"),
    locations = cells_body(
      columns = `KS(aghq, tmbstan)`,
      rows = `KS(aghq, tmbstan)` < `KS(adam, tmbstan)` & `KS(aghq, tmbstan)` < `KS(TMB, tmbstan)`
    )
  ) %>%
    gt::tab_style(
    style = cell_fill(color = "#009E73"),
    locations = cells_body(
      columns = `KS(TMB, tmbstan)`,
      rows = `KS(TMB, tmbstan)` < `KS(aghq, tmbstan)` & `KS(TMB, tmbstan)` < `KS(adam, tmbstan)`
    )
  )
```

This plot could be improved by making the red to green transition smooth, then adding legend for KS difference.
The density plots could be filled according to KS difference.

```{r}
summary_ks_plot <- function(ks_summary, method1, method2) {
  ks_method1 <- paste0("KS(", method1, ", tmbstan)")
  ks_method2 <- paste0("KS(", method2, ", tmbstan)")

  densityplot <- ggplot(ks_summary, aes(x = .data[[ks_method1]], y = .data[[ks_method2]])) +
    stat_density_2d(aes(linetype = Type), col = "black") +
    annotate(geom = "polygon", x = c(-Inf, Inf, Inf), y = c(-Inf, Inf, -Inf), fill = multi.utils::cbpalette()[2], alpha = 0.2) +
    annotate(geom = "polygon", x = c(-Inf, Inf, -Inf), y = c(-Inf, Inf, Inf), fill = multi.utils::cbpalette()[6], alpha = 0.2) +
    xlim(0, 0.75) +
    ylim(0, 0.75) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", alpha = 0.5) +
    scale_linetype_manual(values = c("dashed", "solid")) +
    # scale_alpha_discrete(range = c(0.5, 0.01)) +
    # guides(alpha = FALSE) +
    guides(linetype = FALSE) +
    labs(x = ks_method1, y = ks_method2) +
    theme_minimal()

@@

  ridgeplot <- ggplot(ks_summary, aes(y = Type, x = `KS difference`)) +
    ggridges::geom_density_ridges(alpha = 0.7, fill = NA, aes(linetype = Type)) +
    coord_flip() +
    scale_linetype_manual(values = c("dashed", "solid")) +
    labs(y = "", x = paste0(ks_method1, " - ", ks_method2)) +
    guides(linetype = FALSE) +
    theme_minimal()

  densityplot + ridgeplot
}
